
# Steps Monitor


### Project Overview

The data is number of steps taking in a 5-minute interval across 2 months. The steps below seeks to answer some basic questions about this data.


### Loading and Preprocesing
```{r}
input_data <- read.csv('activity.csv', header=TRUE, stringsAsFactors=FALSE)
input_data$date <- strptime(input_data$date, '%Y-%m-%d')
```


### Mean total number of steps taken per day
```{r steps_per_day}
steps_per_day <- with(input_data, tapply(steps, as.factor(date), sum, na.rm=TRUE))

hist(steps_per_day, col='gray', labels=TRUE)

mean(steps_per_day, na.rm=TRUE)
median(steps_per_day, na.rm=TRUE)
```


### Average daily activity pattern
```{r steps_per_interval}
steps_per_interval <- with(input_data, tapply(steps, as.factor(interval), mean, na.rm=TRUE))
steps_per_interval <- data.frame(interval = as.numeric(names(steps_per_interval)), steps = steps_per_interval)

with(steps_per_interval, plot(interval, steps, type='l'))

max_interval <- steps_per_interval$interval[steps_per_interval$steps == max(steps_per_interval$steps)]
```
The interval with highest average number of steps is `r max_interval`.


### Imputing missing values
```{r steps_per_day_imputed}
sum(is.na(input_data$steps))

imputed_data <- merge(input_data, steps_per_interval, by.x = 'interval', by.y = 'interval')
imputed_data$steps_imputed <- with(imputed_data, ifelse(is.na(steps.x), steps.y, steps.x))
steps_per_day_imputed <- with(imputed_data, tapply(steps_imputed, as.factor(date), sum))

hist(steps_per_day_imputed, col='gray', labels=TRUE)

mean(steps_per_day_imputed, na.rm=TRUE)
median(steps_per_day_imputed, na.rm=TRUE)
```
The mean and median of the imputed steps per day is higher than the mean and median of the steps per day with NA removed.


### Activity patterns between weekdays vs weekends
```{r steps_per_interval_weekday steps_per_interval_weekend}
imputed_data$weekday_type <- ifelse( weekdays(imputed_data$date) %in% c('Saturday', 'Sunday'), 'Weekend', 'Weekday')
data_weekday <- imputed_data[imputed_data$weekday_type == 'Weekday',]
data_weekend <- imputed_data[imputed_data$weekday_type == 'Weekend',]

steps_per_interval_weekday <- with(data_weekday, tapply(steps_imputed, as.factor(interval), sum))
steps_per_interval_weekday <- data.frame(interval = as.numeric(names(steps_per_interval_weekday)), steps = steps_per_interval_weekday)

steps_per_interval_weekend <- with(data_weekend, tapply(steps_imputed, as.factor(interval), sum))
steps_per_interval_weekend <- data.frame(interval = as.numeric(names(steps_per_interval_weekend)), steps = steps_per_interval_weekend)


par(mfrow=c(2,1), mar=c(2,2.5,1,1))
with(steps_per_interval_weekday, plot(interval, steps, type='l', main = 'weekday', xlab=''))
with(steps_per_interval_weekend, plot(interval, steps, type='l', main = 'weekend'))
```
